---
title: "GrCa HRI Exploratory Data Analysis"
author: "Dom Grisafe"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: FALSE
    toc: true
    toc_float: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tableone) # summary table
library(kableExtra) # html tables
library(leaflet) # mapping
```

# Load Analytic Dataset

```{r}
df_screen <- readRDS(paste0(dir_data, "\\df_screen.rds"))
```

# Case Timelines

```{r}
df_screen$dispatch_date %>% range(na.rm = TRUE)
table(lubridate::month(df_screen$dispatch_date, label = TRUE), df_screen$dispatch_year)
```
Missing data from:

* 2016 May to Aug
* 2017 July to Dec
* 2018 Sept to Dec

## Overall Study Period

```{r}
# limit the overall study period
date_lims <- as.POSIXct(strptime(c("2009-01-01", "2023-01-01"), format = "%Y-%m-%d"))    

# plot over the entire study period
df_screen %>% 
  ggplot(aes(x = dispatch_date, fill = lubridate::month(dispatch_date, label = TRUE))) +
  geom_histogram(bins = 13, na.rm = F) +
  scale_x_datetime(
    labels = scales::date_format("'%y"), 
    date_breaks = "1 year",
    limits = date_lims
    ) +
  theme(
    legend.title = element_blank()
  )
```

## Annual Trends

```{r}

```

# Map 1: Case Origins

## Country

## State


# Table 1: Cohort Description{.tabset}

Simple tables for variables in the analytic dataset

## Overall

```{r}
tableone::CreateTableOne(
  data = df_screen, 
  includeNA = FALSE,
  vars = c(
    "age_disp", "gender", "race_ethnicity", "continent", "state", # var_demographic
    "height", "weight", # var_medical
    "dispatch_mos", "dispatch_year" # var_ems
    )
  ) %>% 
  kableone() %>% 
  kable_styling(bootstrap_options = c("striped", "hover")) %>% 
  pack_rows("Demographics", 2, 11) %>% 
  pack_rows("Continent of Origin", 12, 17) %>% 
  pack_rows("State/Providence", 18, 82) %>% 
  pack_rows("Clinical Information", 83, 84) %>% 
  pack_rows("Dispatch Day", 85, 86)
```

## Overall, Missing

```{r}
tableone::CreateTableOne(
  data = df_screen, 
  includeNA = TRUE,
  vars = c(
    "age_disp", "gender", "race_ethnicity", "continent", "state", # var_demographic
    "height", "weight", # var_medical
    "dispatch_mos", "dispatch_year" # var_ems
    )
  ) %>% 
  kableone() %>% 
  kable_styling(bootstrap_options = c("striped", "hover")) %>% 
  pack_rows("Demographics", 2, 13) %>% 
  pack_rows("Continent of Origin", 14, 20) %>% 
  pack_rows("State/Providence", 21, 86) %>% 
  pack_rows("Clinical Information", 87, 88) %>% 
  pack_rows("Dispatch Day", 89, 90)
```

## By Case Status

```{r}
tableone::CreateTableOne(
  data = df_screen, 
  strata = "case_screen", includeNA = TRUE,
  vars = c(
    "age_disp", "gender", "race_ethnicity", "continent", "state", # var_demographic
    "height", "weight", # var_medical
    "dispatch_mos", "dispatch_year" # var_ems
    )
  ) %>% 
  kableone() %>% 
  kable_styling(bootstrap_options = c("striped", "hover")) %>% 
  pack_rows("Demographics", 2, 13) %>% 
  pack_rows("Continent of Origin", 14, 20) %>% 
  pack_rows("State/Providence", 21, 86) %>% 
  pack_rows("Clinical Information", 87, 88) %>% 
  pack_rows("Dispatch Day", 89, 90)
```


# Map 2: Grand Canyon Case Location

```{r}
# create data with coordinates of bases for mapping
base_coord <- tibble(
  base_region = df_screen$base_region %>% levels,
  base_lat  = c(36.056198, 36.1058432, 36.0884229, 36.0029983),
  base_long = c(-112.125198, -112.0968523, -112.1382771, -114.0663743)
)

# add base location data to data
df_grca_map <- merge(base_coord, df_screen, by = "base_region")
```

```{r}
# subset relevant variables for mapping
df_map <- df_grca_map %>% 
  # filter(
  #   base_region != "South Rim"
  #   !is.na(scene_lat) & scene_lat > 35.9 & scene_lat < 36.25 &
  #   !is.na(scene_long) & scene_long > -113 & scene_long < -111 &
  #   !is.na(base_region)
  #   ) %>%
  select(prid, scene_lat, scene_long, 
         base_region, base_lat, base_long
         ) 
```

`r nrow(df_grca_map)` (`r round(100 * nrow(df_grca_map) / nrow(df_screen), digits = 2)`%) cases have latitude and longitude data available

```{r}
# color palette for grouping variable, base region
factpal <- colorFactor(palette = c("#bf9bd9", "#50a9a7", "#8b6464", "#e11404"), df_map$base_region)

# map of regions
leaflet() %>% 
  addProviderTiles(providers$OpenTopoMap, group = "Topo") %>% 
  setView(lng = -113.079, lat = 36.0, zoom = 8) %>% 
  
  # larger circles for base sites
  addCircleMarkers(data = base_coord, lng=~base_long, lat=~base_lat,
                   fillColor = ~factpal(base_region),
                   radius = 40, label = ~base_region, fillOpacity = 0.5, 
                   stroke = TRUE, color = "black", weight = 1
                   ) %>% 
  
  # smaller circles for cases
  addCircleMarkers(data = df_map, lng=~scene_long, lat=~scene_lat,
                   fillColor = ~factpal(base_region), popup = ~prid,
                   radius = 5, label = ~base_region, fillOpacity = 10,
                   stroke = TRUE, color = "white", weight = 1
                   ) %>% 
  
  # legend settings
  addLegend('bottomright', pal = factpal, values = df_map$base_region,
            title = 'Base Region', opacity = 5
            )
```


# Table 2: Medicial Conditions, Medications

# Table 3: Activity Engaged

# Table 4: Paramedic Assessments

# Figure 1: Temperature and Rate per 100,000 on Month

Where do we get the temperature data?

Should we try a three dimensional plot where the z-axis is the year? This would consolidate many graphs, but may be too confusing to interpret.
