---
title: "GrCa HRI Exploratory Data Analysis"
author: "Dom Grisafe"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: FALSE
    toc: true
    toc_float: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # hide code from html output
library(tidyverse)
library(tableone) # summary table
library(kableExtra) # html tables
library(leaflet) # mapping
library(lubridate)
library(rgl) # 3D plotting
```

# To-Do List

* [ ] Set unrealistic latitude and longitude case data as missing, visualize with GrCa case map
* [ ] Get missing data from Maria in SharePoint
    * 2016 May to Aug
    * 2017 July to Dec
    * 2018 Sept to Dec
* [ ] Add temperature data using [getGridMet](https://rdrr.io/github/mikejohnson51/climateR/man/getGridMET.html) 
* [ ] Get park visitor denominator data from Danielle, who's colleague works next to her

# Load Analytic Dataset

```{r}
df_screen <- readRDS(paste0(dir_data, "\\df_screen.rds"))
```

# Case Timelines{.tabset}

Check data to ensure no missing from various time periods.

```{r}
table(df_screen$dispatch_year, lubridate::month(df_screen$dispatch_date, label = TRUE))
```

Missing data from:

* 2016 May to Aug
* 2017 July to Dec
* 2018 Sept to Dec

## Overall Study Period

```{r}
# limit the overall study period
date_lims <- as.POSIXct(strptime(c("2009-01-01", "2023-01-01"), format = "%Y-%m-%d"))    

# plot over the entire study period
df_screen %>% 
  ggplot(aes(x = dispatch_date, fill = lubridate::month(dispatch_date, label = TRUE))) +
  geom_histogram(bins = 13, na.rm = T) +
  scale_x_datetime(
    labels = scales::date_format("'%y"), 
    date_breaks = "1 year",
    limits = date_lims
    ) +
  theme(
    legend.title = element_blank()
  )
```

## Overall Study Period by Year

```{r}
# limit the overall study period
date_lims <- as.POSIXct(strptime(c("2009-01-01", "2023-01-01"), format = "%Y-%m-%d"))    

# plot over the entire study period
df_screen %>% 
  ggplot(aes(x = dispatch_date, fill = lubridate::month(dispatch_date, label = TRUE))) +
  geom_histogram(bins = 12, na.rm = T) +
  scale_x_datetime(
    labels = scales::date_format("%m"), 
    date_breaks = "2 months"
    ) +
  facet_wrap(~dispatch_year, scales ='free_x') +
  theme(
    legend.title = element_blank(),
    legend.position = "bottom",
  ) +
  guides(
    fill = guide_legend(label.position = "right", nrow = 1), 
    )
```

## Annual Trends

```{r}
df_screen %>% 
  ggplot(aes(yday(dispatch_date), group=factor(year(dispatch_date)), fill=factor(year(dispatch_date)))) + 
  geom_histogram(bins = 12, na.rm = T) +
  theme(
    legend.title = element_blank(),
    legend.position = "right",
  ) +
  guides(
    fill = guide_legend(label.position = "right", ncol = 1),
    )
```

## Monthly Trends

```{r}
df_monthly <- df_screen %>% 
  group_by(dispatch_year, dispatch_mos) %>% 
  summarise(count = n()) %>% 
  mutate(dispatch_year_f = factor(dispatch_year)) %>% 
  ungroup()

df_monthly %>% 
  ggplot(aes(x = dispatch_mos, y = count, linetype = dispatch_year_f, color = dispatch_year_f, scale = 1/10000)) +
    geom_line(na.rm=T) +
  scale_x_continuous(breaks = seq(1,12,1)) +
  ylim(0,175) +
  theme(
    legend.title = element_blank(),
    legend.position = "right",
  ) +
  guides(
    fill = guide_legend(label.position = "right", ncol = 1),
    )
```

## 3D Trends

3D plotting using the [`rgl`](https://cran.r-project.org/web/packages/rgl/vignettes/rgl.html) library package.

```{r}
# clear existing 3d html window
clear3d()

# output rgl plot in a rmarkdown document
setupKnitr()

# saved coordinates from view I like
myview <- matrix(nrow = 4, ncol = 4, byrow = TRUE,
                 data = c(0.8436018, 0.5357571, -0.03606263,  0,
                          -0.3189417, 0.5539688, 0.76902199,  0,
                           0.4319866, -0.6372465, 0.63820428,  0,
                           0,          0,          0,           1
                          )
                 )

# # open 3d window
# open3d()

# Add a new column with color
my_col_pal <- scales::col_numeric(domain = c(0, 168), palette = "Spectral", reverse = TRUE)
df_monthly$color <- my_col_pal(df_monthly$count)


## triangulate a surface (https://stackoverflow.com/questions/53398108/how-to-plot-a-surface-in-rgl-plot3d)
# remove missing row prevents error
df_monthly <- df_monthly %>% filter(!is.na(dispatch_mos) & !is.na(dispatch_year)) 
# triangulate data
dxyz <- deldir::deldir(x = df_monthly$dispatch_year, y = df_monthly$dispatch_mos, z = df_monthly$count)
# 3D plot in rgl
persp3d(dxyz, col = df_monthly$color, 
        xlim=c(2009,2022), ylim = c(0,12), zlim = c(0, 170),
        aspect = c(4,2,1),# zoom on z axis
        xlab="Year", ylab="Month", zlab="Cases"
        )


# # save orientation of 3d object (https://cran.r-project.org/web/packages/rgl/vignettes/rgl.html#controlling-the-look-of-the-scene)
# myview2 <- par3d("userMatrix")

# call coordinates
par3d(userMatrix = myview)

# Add bounding box decoration
bbox3d(color = c("#ffffff", "black"), alpha = 0.2)

# To display in an R Markdown document:
rglwidget()

# # close 3d window
# close3d()

# # To save to a file:
# htmlwidgets::saveWidget(rglwidget(width = 520, height = 520),
#                         file = "HtmlWidget-3dscatter.html",
#                         libdir = "libs",
#                         selfcontained = FALSE
#                         )
```


# Map 1: Case Origins

## Country

## State


# Table 1: Cohort Description{.tabset}

Simple tables for variables in the analytic dataset

## Overall

```{r}
tableone::CreateTableOne(
  data = df_screen, 
  includeNA = FALSE,
  vars = c(
    "age_disp", "gender", "race_ethnicity", "continent", "state", # var_demographic
    "height", "weight", # var_medical
    "dispatch_mos", "dispatch_year" # var_ems
    )
  ) %>% 
  kableone() %>% 
  kable_styling(bootstrap_options = c("striped", "hover")) %>% 
  pack_rows("Demographics", 2, 11) %>% 
  pack_rows("Continent of Origin", 12, 17) %>% 
  pack_rows("State/Providence", 18, 82) %>% 
  pack_rows("Clinical Information", 83, 84) %>% 
  pack_rows("Dispatch Day", 85, 86)
```

## Overall, Missing

```{r}
tableone::CreateTableOne(
  data = df_screen, 
  includeNA = TRUE,
  vars = c(
    "age_disp", "gender", "race_ethnicity", "continent", "state", # var_demographic
    "height", "weight", # var_medical
    "dispatch_mos", "dispatch_year" # var_ems
    )
  ) %>% 
  kableone() %>% 
  kable_styling(bootstrap_options = c("striped", "hover")) %>% 
  pack_rows("Demographics", 2, 13) %>% 
  pack_rows("Continent of Origin", 14, 20) %>% 
  pack_rows("State/Providence", 21, 86) %>% 
  pack_rows("Clinical Information", 87, 88) %>% 
  pack_rows("Dispatch Day", 89, 90)
```

## By Case Status

```{r}
tableone::CreateTableOne(
  data = df_screen, 
  strata = "case_screen", includeNA = TRUE,
  vars = c(
    "age_disp", "gender", "race_ethnicity", "continent", "state", # var_demographic
    "height", "weight", # var_medical
    "dispatch_mos", "dispatch_year" # var_ems
    )
  ) %>% 
  kableone() %>% 
  kable_styling(bootstrap_options = c("striped", "hover")) %>% 
  pack_rows("Demographics", 2, 13) %>% 
  pack_rows("Continent of Origin", 14, 20) %>% 
  pack_rows("State/Providence", 21, 86) %>% 
  pack_rows("Clinical Information", 87, 88) %>% 
  pack_rows("Dispatch Day", 89, 90)
```


# Map 2: Grand Canyon Case Location

```{r}
# create data with coordinates of bases for mapping
base_coord <- tibble(
  base_region = df_screen$base_region %>% levels,
  base_lat  = c(36.056198, 36.1058432, 36.0884229, 36.0029983),
  base_long = c(-112.125198, -112.0968523, -112.1382771, -114.0663743)
)

# add base location data to data
df_grca_map <- merge(base_coord, df_screen, by = "base_region")
```

```{r}
# subset relevant variables for mapping
df_map <- df_grca_map %>% 
  # filter(
  #   base_region != "South Rim"
  #   !is.na(scene_lat) & scene_lat > 35.9 & scene_lat < 36.25 &
  #   !is.na(scene_long) & scene_long > -113 & scene_long < -111 &
  #   !is.na(base_region)
  #   ) %>%
  select(prid, scene_lat, scene_long, 
         base_region, base_lat, base_long
         ) 
```

`r nrow(df_grca_map)` (`r round(100 * nrow(df_grca_map) / nrow(df_screen), digits = 2)`%) cases have latitude and longitude data available

```{r}
# color palette for grouping variable, base region
factpal <- colorFactor(palette = c("#bf9bd9", "#50a9a7", "#8b6464", "#e11404"), df_map$base_region)

# map of regions
leaflet() %>% 
  addProviderTiles(providers$OpenTopoMap, group = "Topo") %>% 
  setView(lng = -113.079, lat = 36.0, zoom = 8) %>% 
  
  # larger circles for base sites
  addCircleMarkers(data = base_coord, lng=~base_long, lat=~base_lat,
                   fillColor = ~factpal(base_region),
                   radius = 40, label = ~base_region, fillOpacity = 0.5, 
                   stroke = TRUE, color = "black", weight = 1
                   ) %>% 
  
  # smaller circles for cases
  addCircleMarkers(data = df_map, lng=~scene_long, lat=~scene_lat,
                   fillColor = ~factpal(base_region), popup = ~prid,
                   radius = 5, label = ~base_region, fillOpacity = 10,
                   stroke = TRUE, color = "white", weight = 1
                   ) %>% 
  
  # legend settings
  addLegend('bottomright', pal = factpal, values = df_map$base_region,
            title = 'Base Region', opacity = 5
            )
```


# Table 2: Medicial Conditions, Medications

```{r}
# get the number of medications any given person is taking, create new column
df_screen$n_meds <- df_screen$medications %>% str_split(pattern = ", ") %>% lapply(length) %>% unlist
n_max_meds <- df_screen$n_meds %>% max

# create dataframe of medications in long format, multiple rows per person if taking multiple medications
df_meds_long <- df_screen %>% 
  # separate medications delimited by a comma and single white space
  tidyr::separate(col = medications, sep = ", ", into = paste0("medication_", 1:n_max_meds), remove = FALSE, fill = "right") %>% 
  # show only patient id and medications
  dplyr::select(prid, starts_with("medication")) %>% 
  # pivot from wide to long data, name new columns, drop places when patient not taking multiple medications
  pivot_longer(cols = starts_with("medication_"), names_to = "n_med", values_to = "medication", values_drop_na = TRUE) 
  
# table of frequency counts
t2_meds <- df_meds_long$medication %>% table() %>%
  # sort in descending order
  sort(decreasing = TRUE)

# create data frame of medications by frequency
df_t2_meds <- tibble(
  Medication = names(t2_meds),
  Frequency = t2_meds
)

# create HTML table of medications by frequency
df_t2_meds %>% 
  kable("html", align = c("l", "r")) %>% 
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```


# Table 3: Activity Engaged

# Table 4: Paramedic Assessments

# Figure 1: Temperature and Rate per 100,000 on Month

Where do we get the temperature data?

Should we try a three dimensional plot where the z-axis is the year? This would consolidate many graphs, but may be too confusing to interpret.
