---
title: "HRI Grand Canyon"
author: "Dom Grisafe"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: FALSE
    toc: true
    toc_float: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
```

```{r, echo=FALSE}
details::details(object = sessioninfo::session_info(), summary = "Current Session Info")
```


# Data Cleaning




## Load Data

Load data from .csv file

Format missing data as NA from various missing label entries (e.g., "Not Available", "Not Known", etc.)

```{r}
# load data from .csv file
data_csv <- read_csv(
  paste0(dir_data, "\\20230206_GRCA_2021A.csv"),
  # quiet warning
  show_col_types = FALSE,
  
  # replace missing string indicators as NA
  na = c("", "Not Available", "Not Known", "Unknown", "Not Applicable", "unknown")
  )
```


## Format Variable Names 

Format variable names as snake_case

```{r}
# clean variable names to snake_case
names(data_csv) <- data_csv %>% names() %>% 
  # make alll variable names lower case
  str_to_lower() %>% 
  # remove punctuation from all variable names
  str_remove_all("[[:punct:]]+") %>% 
  # make any white space in variable names into a single underscore
  str_replace_all("\\s+","_")
```


## Format Existing Variables

Existing variables may be:

* formatted as factors, characters, dates, etc.
* renamed (see *Remove Irrelevant Variables* for excluding original names)
* separated into smaller variables from complex variables

```{r}
# clean existing variables
df_format <- data_csv %>% 
  
  # clean variables and format
  mutate(
    
    # convert clean categorical variables from string to factor
    medical_category = factor(medical_category),
    patient_activity = factor(patient_activity),
    gender = factor(gender),
    service_name = factor(service_name),
    type_of_service_ihscene = factor(type_of_service_ihscene),
    medical_category_historical = factor(medical_category_historical),
    
    # format as dates and time for dates of dispatch and birth
    dispatch_date = as.POSIXct(date_dispatched, format = "%m/%d/%Y %H:%M"),
    date_of_birth = as.POSIXct(date_of_birth, format = "%m/%d/%Y %H:%M"),
    
    # create dispatch time variables for date, mos, year
    dispatch_day = lubridate::day(dispatch_date),
    dispatch_mos = lubridate::month(dispatch_date),
    dispatch_year = lubridate::year(dispatch_date),
    
    # make id variable a character
    prid = as.character(prid),
    
    # format base region within park
    base_region = case_when(
      str_detect(basesite, "South Rim") ~ 1,
      str_detect(basesite, "Phantom Ranch") ~ 2,
      str_detect(basesite, "Indian Garden") ~ 3,
      str_detect(basesite, "Meadview") ~ 4
    ) %>% factor(levels = 1:4, labels = c("South Rim", "Phantom Ranch", "Indian Garden", "Meadview")),
    
    # format base type
    base_type = case_when(
      str_detect(basesite, "Station 1- South Rim") ~ 1,
      str_detect(basesite, "South Rim Helibase") ~ 2,
      str_detect(basesite, "South Rim- Ranger Operations") ~ 3,
      str_detect(basesite, "Phantom Ranch Ranger Station") ~ 4,
      str_detect(basesite, "Indian Garden Ranger Station") ~ 4,
      str_detect(basesite, "Meadview Ranger Station") ~ 4
      ) %>% factor(levels = 1:4, labels = c("Station 1", "Helibase", "Ranger Operations", "Ranger Station")),
    
    # format race/ethnicity
    race_ethnicity = factor(race, ordered = TRUE,c(
      "American Indian / Alaska Native", "Asian / Pacific Islander", 
      "White, Hispanic", "White, Non-Hispanic"
      )),
    
    # format response priority
    dispatch_priority = factor(dispatch_priority, ordered = TRUE,
      levels = c("Alpha Response", "Beta Response", "Charlie Response", "Delta Response")
    ),
    
    # format alcohol from alcohol_or_drugs
    alcohol = case_when(
      str_detect(alcohol_or_drugs, "Alcohol") ~ 1,
      str_detect(alcohol_or_drugs, "None") ~ 2
    ) %>% form_yn(),
    
    # format drugs from alcohol_or_drugs
    drugs = case_when(
      str_detect(alcohol_or_drugs, "Drugs") ~ 1,
      str_detect(alcohol_or_drugs, "None") ~ 2
    ) %>% form_yn(),
    
    # format outcome variable as dichotomous treated yes/no
    outcome_treated = case_when(
      str_detect(outcome_historical_all, "Treated") ~ 1,
      !str_detect(outcome_historical_all, "Treated") ~ 2
    ) %>% form_yn(),
    
    # format first agency unit on scene
    first_on_scene = case_when(
      first_agency_unit_on_scene == "Yes" ~ 1,
      first_agency_unit_on_scene == "No" ~ 2,
      first_agency_unit_on_scene == "Not Applicable" ~ 3
      ) %>% factor(1:3, c("Yes", "No", "Not Applicable")),
    
    # format park unit
    unit = factor(unit, ordered = TRUE, 
      labels = c("Helo 8PA", "Helo 368", "Medic 81", "Medic 82", "Medic 84", "Park Ranger")
      ),
    
    # format response mode
    response_mode = factor(response_mode, ordered = T,
      labels = c("Emergent (Immediate Response)", "Non-Emergent", "No Lights/Sirens")
    ),    
    
    # format transport mode
    transport_mode = factor(transport_mode, ordered = T,
      labels = c("Emergent (Immediate Response)", "Non-Emergent", "No Lights/Sirens")
    ),
    
    # format address variables
    address_line_1 = str_to_title(address_line_1) %>% str_replace("Po |P.o. ", "PO "),
    address_line_2 = str_to_title(address_line_2) %>% str_replace("Po |P.o. ", "PO "),
    city = str_to_title(city),
    state = str_to_title(state),
    country = str_to_title(country),
    
    # name variables renamed and formatted as titles
    name_last = str_to_title(last_name),
    name_first = str_to_title(first_name),
    name_middle = str_to_title(middle_name),
    
    # phone in consistent format
    # from [R4DS](https://r4ds.had.co.nz/strings.html?q=phone#other-types-of-pattern)
    phone = str_extract(phone, 
      pattern = regex(ignore_case = TRUE, comments = TRUE,
        "
          \\(?     # optional opening parens
          (\\d{3}) # area code
          [) -]?   # optional closing parens, space, or dash
          (\\d{3}) # another three numbers
          [ -]?    # optional space or dash
          (\\d{4}) # four more numbers
        "
      )
    ),
    
    # medications patients taking
    medications = 
      # set common indicator variable for no medications
      ifelse(
        test = str_detect(medications, pattern = regex('^none|denied|denies', ignore_case = T)),
        yes = "None", 
        no = ifelse(
          # set NA for unknown whether taking medications
          test = str_detect(medications, pattern = regex('^unknown', ignore_case = T)),
          yes = NA, 
          no = medications
          )
        ) %>% 
      # replace unk prefix with complete word unknown
      str_replace_all(regex("^unk", ignore_case = TRUE), "unknown") %>% 
      # add spaces after commas separating drug names
      # https://stackoverflow.com/questions/44155594/add-a-space-after-commas-in-a-string-only-if-it-doesnt-exist-in-r
      str_replace_all(pattern = ",([[:alpha:]])", replacement = ", \\1") %>% 
      # replace and with comma, consistent with formatting
      str_replace_all(pattern = " and ", replacement = ", ") %>% 
      # Capitalize Each Word
      str_to_title(),
    
    # allergies to medications
    allergies = 
      # set common indicator variable for no known allergies
      ifelse(
        test = str_detect(allergies, pattern = regex('^none|denied|denies|nkda|nka', ignore_case = T)),
        yes = "No Known Allergies", 
        no = ifelse(
          # set NA for unknown whether have allergies
          test = str_detect(allergies, pattern = regex('^unknown', ignore_case = T)),
          yes = NA, 
          no = allergies
          )
        ) %>% 
      # add spaces after commas separating drug names
      # https://stackoverflow.com/questions/44155594/add-a-space-after-commas-in-a-string-only-if-it-doesnt-exist-in-r
      str_replace_all(pattern = ",([[:alpha:]])", replacement = ", \\1") %>% 
      # replace and with comma, consistent with formatting
      str_replace_all(pattern = " and ", replacement = ", ") %>% 
      # Capitalize Each Word
      str_to_title()
  )
```


## Derived Variables

Variables that are created for later analysis from existing variables

```{r}
# derived variables
df_derived <- df_format %>% 
  
  mutate(
    
    # age at date of dispatch from date of birth
    age_disp = as.numeric(interval(date_of_birth, dispatch_date),"years"),
    
    # add missing country as "United States" if state is in names of USA States
    country = ifelse(state %in% state.name & is.na(country), "United States", country),
    
    # create composite address variable with street address, city, state (if US or Canada), country
    address = paste(sep = ", ",
      # combine first and second street addresses, if the second address exists
      ifelse(!is.na(address_line_2), paste(sep = ", ", address_line_1, address_line_2), address_line_1),
      # if address in USA or Canada, then combine city, state, country
      # else combine city, country
      ifelse(country %in% c("United States", "Canada"), paste(city, state, country, sep = ", "), paste(city, country, sep = ", "))
      ),  
    
    # create full name variable
    name_full = ifelse(
      !is.na(name_middle), 
      yes = paste(name_last, paste(name_first, name_middle), sep = ", "),
      no = paste(name_last, name_first, sep = ", ")
    )
    
  ) %>% 
  
  # split scene location gps coordinates into separate variables for latitude and longitude
  extract(
    scene_location_gps, c("scene_lat", "scene_long"), 
    "([0-9]+\\.[0-9]+)\\,(\\-[0-9]+\\.[0-9]+)"
    )
```


## Remove Irrelevant Variables

Remove variables that are irrelevant because they were:

* used in derived variables and are now redundant
* not relevant to this analysis
* all values were missin; note, this may change when longer data sets are added
* duplicate variables that are redundant within dataset

```{r}
# remove irrelevant variables
df_concise <- df_derived %>% select(
  
  # remove variables used to make cleaner variables above in mutate() function 
  -basesite, -date_dispatched, -day_of_month_of_dispatch, -month_of_dispatch_alltime,
  -year_of_dispatch, -alcohol_or_drugs, -first_agency_unit_on_scene, -age,
  -age_in_years_calc, -last_name, -first_name, -middle_name, -race,

  # remove unimportant variables
  -receiving_other, -receiving_ems, -receiving_hospital, -receiving_hospital_designation,
  -receiving_hospital_unit, -receiving_md, -receiving_mode_descriptors, 
  -receiving_mode_descriptors_list, -receiving_rn, -receiving_type, -receiving_usng,
  -anatomic_location,

  # remove variables with all missing data
  # note, this may change as more observations are added to future datasets
  # therefore reassess these variables when adding data
  -impression_primary_nemsis_2, -other_agencies,
  -other_ems_name, -impression_list, 
  -primary_diagnosis, -impression, -secondary_complaint, -working_diagnosis,

  # remove duplicate variables
  -disposition_outcome # outcome_historical_all / outcome_treated
  
  )
```

## Rename Variables by Categorizing into Groups

Some variables naturally group together.

I'll add prefixes to each group of variables:

* Identification (ide_)
* Demographic (dem_)
* Subjective (sub_)
* Emergency Medical Services (ems_)

```{r}

```

## Manual Entries

Some calculated variables are missing because one of the input variables is missing.

Sometimes duplicate variables exist for the calculated variables. We'll use these here to substitute for the missing values.

```{r}
# manually enter values for missing data
df_concise[df_concise$prid=="66267999","age_disp"] <- 58 # taken from var age_in_years_calc
```


## Duplicate Entries

```{r}
df_concise %>% dim()
df_concise %>% select(-patient_activity) %>% duplicated() %>% sum()
df_concise %>% select(-patient_activity) %>% distinct %>% dim()
df_grca <- df_concise
```

It looks like there are 12 entries that are duplicated except by the variable  `patient_activity`.

This is long data that needs to be made wide as `patient_activity_1` and `patient_activity_2`

## View Final Data

I've completed up until the variable `phone`.

I have variables to finish cleaning up through `working_diagnosis` and before `dispatch_date`.

```{r}
df_grca2 <- df_grca %>% mutate(

)
```

```{r}
# df_grca %>% mutate(
# 
# ) %>% select(address, address_line_1, address_line_2, city, state, country) %>% unique
# test <- tibble(
#   df_grca2$medications,
#   data_csv$medications
# )
# df_grca %>% unique()
# df_grca %>% glimpse
```


# Data Dictionary

This will be the data dictionary.


# Analytic Data

Simple tables for variables in the analytic dataset
