---
title: "HRI Grand Canyon"
author: "Dom Grisafe"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: FALSE
    toc: true
    toc_float: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
```

```{r, echo=FALSE}
details::details(object = sessioninfo::session_info(), summary = "Current Session Info")
```


# Data Cleaning




## Load Data

Load data from .csv file

Format missing data as NA from various missing label entries (e.g., "Not Available", "Not Known", etc.)

```{r}
# load data from .csv file
data_csv <- read_csv(
  paste0(dir_data, "\\20230206_GRCA_2021A.csv"),
  # quiet warning
  show_col_types = FALSE,
  
  # replace missing string indicators as NA
  na = c("", "Not Available", "Not Known", "Unknown", "Not Applicable", "unknown")
  )
```


## Format Variable Names 

Format variable names as snake_case

```{r}
# clean variable names to snake_case
names(data_csv) <- data_csv %>% names() %>% 
  # make alll variable names lower case
  str_to_lower() %>% 
  # remove punctuation from all variable names
  str_remove_all("[[:punct:]]+") %>% 
  # make any white space in variable names into a single underscore
  str_replace_all("\\s+","_")
```


## Format Existing Variables

Existing variables may be:

* formatted as factors, characters, dates, etc.
* renamed (see *Remove Irrelevant Variables* for excluding original names)
* separated into smaller variables from complex variables

```{r}
# clean existing variables
df_format <- data_csv %>% 
  
  # clean variables and format
  mutate(
    
    # convert clean categorical variables from string to factor
    medical_category = factor(medical_category),
    patient_activity = factor(patient_activity),
    gender = factor(gender),
    service_name = factor(service_name),
    type_of_service_ihscene = factor(type_of_service_ihscene),
    medical_category_historical = factor(medical_category_historical),
    
    # format as dates and time for dates of dispatch and birth
    dispatch_date = as.POSIXct(date_dispatched, format = "%m/%d/%Y %H:%M"),
    date_of_birth = as.POSIXct(date_of_birth, format = "%m/%d/%Y %H:%M"),
    
    # create dispatch time variables for date, mos, year
    dispatch_day = lubridate::day(dispatch_date),
    dispatch_mos = lubridate::month(dispatch_date),
    dispatch_year = lubridate::year(dispatch_date),
    
    # make id variable a character
    prid = as.character(prid),
    
    # format base region within park
    base_region = case_when(
      str_detect(basesite, "South Rim") ~ 1,
      str_detect(basesite, "Phantom Ranch") ~ 2,
      str_detect(basesite, "Indian Garden") ~ 3,
      str_detect(basesite, "Meadview") ~ 4
    ) %>% factor(levels = 1:4, labels = c("South Rim", "Phantom Ranch", "Indian Garden", "Meadview")),
    
    # format base type
    base_type = case_when(
      str_detect(basesite, "Station 1- South Rim") ~ 1,
      str_detect(basesite, "South Rim Helibase") ~ 2,
      str_detect(basesite, "South Rim- Ranger Operations") ~ 3,
      str_detect(basesite, "Phantom Ranch Ranger Station") ~ 4,
      str_detect(basesite, "Indian Garden Ranger Station") ~ 4,
      str_detect(basesite, "Meadview Ranger Station") ~ 4
      ) %>% factor(levels = 1:4, labels = c("Station 1", "Helibase", "Ranger Operations", "Ranger Station")),
    
    # format race/ethnicity
    race_ethnicity = factor(race, ordered = TRUE,c(
      "American Indian / Alaska Native", "Asian / Pacific Islander", 
      "White, Hispanic", "White, Non-Hispanic"
      )),
    
    # format response priority
    dispatch_priority = factor(dispatch_priority, ordered = TRUE,
      levels = c("Alpha Response", "Beta Response", "Charlie Response", "Delta Response")
    ),
    
    # format alcohol from alcohol_or_drugs
    alcohol = case_when(
      str_detect(alcohol_or_drugs, "Alcohol") ~ 1,
      str_detect(alcohol_or_drugs, "None") ~ 2
    ) %>% form_yn(),
    
    # format drugs from alcohol_or_drugs
    drugs = case_when(
      str_detect(alcohol_or_drugs, "Drugs") ~ 1,
      str_detect(alcohol_or_drugs, "None") ~ 2
    ) %>% form_yn(),
    
    # format outcome variable as dichotomous treated yes/no
    outcome_treated = case_when(
      str_detect(outcome_historical_all, "Treated") ~ 1,
      !str_detect(outcome_historical_all, "Treated") ~ 2
    ) %>% form_yn(),
    
    # format first agency unit on scene
    first_on_scene = case_when(
      first_agency_unit_on_scene == "Yes" ~ 1,
      first_agency_unit_on_scene == "No" ~ 2,
      first_agency_unit_on_scene == "Not Applicable" ~ 3
      ) %>% factor(1:3, c("Yes", "No", "Not Applicable")),
    
    # format park unit
    unit = factor(unit, ordered = TRUE, 
      labels = c("Helo 8PA", "Helo 368", "Medic 81", "Medic 82", "Medic 84", "Park Ranger")
      ),
    
    # format response mode
    response_mode = factor(response_mode, ordered = T,
      labels = c("Emergent (Immediate Response)", "Non-Emergent", "No Lights/Sirens")
    ),    
    
    # format transport mode
    transport_mode = factor(transport_mode, ordered = T,
      labels = c("Emergent (Immediate Response)", "Non-Emergent", "No Lights/Sirens")
    ),
    
    # format address variables
    address_line_1 = str_to_title(address_line_1) %>% str_replace("Po |P.o. ", "PO "),
    address_line_2 = str_to_title(address_line_2) %>% str_replace("Po |P.o. ", "PO "),
    city = str_to_title(city),
    state = str_to_title(state),
    country = str_to_title(country),
    
    # name variables renamed and formatted as titles
    name_last = str_to_title(last_name),
    name_first = str_to_title(first_name),
    name_middle = str_to_title(middle_name),
    
    # phone in consistent format
    # from [R4DS](https://r4ds.had.co.nz/strings.html?q=phone#other-types-of-pattern)
    phone = str_extract(phone, 
      pattern = regex(ignore_case = TRUE, comments = TRUE,
        "
          \\(?     # optional opening parens
          (\\d{3}) # area code
          [) -]?   # optional closing parens, space, or dash
          (\\d{3}) # another three numbers
          [ -]?    # optional space or dash
          (\\d{4}) # four more numbers
        "
      )
    ),
    
    # medications patients taking
    medications = 
      # set common indicator variable for no medications
      ifelse(
        test = str_detect(medications, pattern = regex('^none|denied|denies', ignore_case = T)),
        yes = "None", 
        no = ifelse(
          # set NA for unknown whether taking medications
          test = str_detect(medications, pattern = regex('^unknown', ignore_case = T)),
          yes = NA, 
          no = medications
          )
        ) %>% 
      # replace unk prefix with complete word unknown
      str_replace_all(regex("^unk", ignore_case = TRUE), "unknown") %>% 
      # add spaces after commas separating drug names
      # https://stackoverflow.com/questions/44155594/add-a-space-after-commas-in-a-string-only-if-it-doesnt-exist-in-r
      str_replace_all(pattern = ",([[:alpha:]])", replacement = ", \\1") %>% 
      # replace and with comma, consistent with formatting
      str_replace_all(pattern = " and ", replacement = ", ") %>% 
      # Capitalize Each Word
      str_to_title(),
    
    # allergies to medications
    allergies = 
      # set common indicator variable for no known allergies
      ifelse(
        test = str_detect(allergies, pattern = regex('^none|denied|denies|nkda|nka', ignore_case = T)),
        yes = "No Known Allergies", 
        no = ifelse(
          # set NA for unknown whether have allergies
          test = str_detect(allergies, pattern = regex('^unknown', ignore_case = T)),
          yes = NA, 
          no = allergies
          )
        ) %>% 
      # add spaces after commas separating drug names
      # https://stackoverflow.com/questions/44155594/add-a-space-after-commas-in-a-string-only-if-it-doesnt-exist-in-r
      str_replace_all(pattern = ",([[:alpha:]])", replacement = ", \\1") %>% 
      # replace and with comma, consistent with formatting
      str_replace_all(pattern = " and ", replacement = ", ") %>% 
      # Capitalize Each Word
      str_to_title()
  )
```


## Derived Variables

Variables that are created for later analysis from existing variables

```{r}
# derived variables
df_derived <- df_format %>% 
  
  mutate(
    
    # age at date of dispatch from date of birth
    age_disp = as.numeric(interval(date_of_birth, dispatch_date),"years"),
    
    # add missing country as "United States" if state is in names of USA States
    country = ifelse(state %in% state.name & is.na(country), "United States", country),
    
    # create composite address variable with street address, city, state (if US or Canada), country
    address = paste(sep = ", ",
      # combine first and second street addresses, if the second address exists
      ifelse(!is.na(address_line_2), paste(sep = ", ", address_line_1, address_line_2), address_line_1),
      # if address in USA or Canada, then combine city, state, country
      # else combine city, country
      ifelse(country %in% c("United States", "Canada"), paste(city, state, country, sep = ", "), paste(city, country, sep = ", "))
      ),  
    
    # create full name variable
    name_full = ifelse(
      !is.na(name_middle), 
      yes = paste(name_last, paste(name_first, name_middle), sep = ", "),
      no = paste(name_last, name_first, sep = ", ")
    )
    
  ) %>% 
  
  # split scene location gps coordinates into separate variables for latitude and longitude
  extract(
    scene_location_gps, c("scene_lat", "scene_long"), 
    "([0-9]+\\.[0-9]+)\\,(\\-[0-9]+\\.[0-9]+)"
    )
```

## Remove Irrelevant Variables

Remove variables that are irrelevant because they were:

* used in derived variables and are now redundant
* not relevant to this analysis
* all values were missin; note, this may change when longer data sets are added
* duplicate variables that are redundant within dataset

```{r}
# remove irrelevant variables
df_concise <- df_derived %>% select(
  
  # remove variables used to make cleaner variables above in mutate() function 
  -basesite, -date_dispatched, -day_of_month_of_dispatch, -month_of_dispatch_alltime,
  -year_of_dispatch, -alcohol_or_drugs, -first_agency_unit_on_scene, -age,
  -age_in_years_calc, -last_name, -first_name, -middle_name, -race,

  # remove unimportant variables
  -receiving_other, -receiving_ems, -receiving_hospital, -receiving_hospital_designation,
  -receiving_hospital_unit, -receiving_md, -receiving_mode_descriptors, 
  -receiving_mode_descriptors_list, -receiving_rn, -receiving_type, -receiving_usng,
  -anatomic_location,

  # remove variables with all missing data
  # note, this may change as more observations are added to future datasets
  # therefore reassess these variables when adding data
  -impression_primary_nemsis_2, -other_agencies,
  -other_ems_name, -impression_list, 
  -primary_diagnosis, -impression, -secondary_complaint, -working_diagnosis,

  # remove duplicate variables
  -disposition_outcome # outcome_historical_all / outcome_treated
  
  )
```


## Manual Entries

Some calculated variables are missing because one of the input variables is missing.

Sometimes duplicate variables exist for the calculated variables. We'll use these here to substitute for the missing values.

```{r}
# manually enter values for missing data
df_concise[df_concise$prid=="66267999","age_disp"] <- 58 # taken from var age_in_years_calc
```


## Duplicate Entries

```{r}
# size/dimensions of entire dataset
df_concise %>% dim()

# patient_activity appears to be the only variable that is unique to every row; other columns repeat and are not unique for all observations/rows
df_concise %>% select(prid, starts_with("patient_activity")) %>% head()

# identify all duplicate rows, assuming only patient_activity is unique
duplicate_rows <- df_concise %>% select(-patient_activity) %>% duplicated() 

# identify patient identifiers of all duplicated observations
test_id <- df_concise[duplicate_rows,]$prid
```

`r sum(duplicate_rows)` duplicate rows/observations after excluding the variable `patient_activity`.

This long data that must be spread as `patient_activity_1`, `patient_activity_2`, etc.

Create unique rows per patient by [grouping, creating a unique ID variable, ungrouping, spreading](https://stackoverflow.com/questions/50117777/multiple-rows-into-multiple-columns-dataframe-in-r)

```{r}
# spread repeating patient_activity data into multiple columns so each row is one unique person
# relearned from https://stackoverflow.com/questions/50117777/multiple-rows-into-multiple-columns-dataframe-in-r
df_unique <- df_concise %>% 
  # select all variables except for the unique variable
  group_by_at(vars(-patient_activity)) %>% 
  # create ID variable that is unique for each patient activity
  mutate(ID = paste0("patient_activity_", 1:n())) %>% 
  ungroup() %>% 
  # spread unique patient activities into separate columns
  spread(ID, patient_activity)
```

## Rename and Order Variables as Groups

Some variables naturally group together.

Add prefixes to each group of variables:

* Identification (id_)
* Demographic (dem_)
* Medical (med_)
* Emergency Medical Services (ems_)

```{r}
# identification variables
var_id <- c(
  "dispatch_id", "prid",
  "name_full", "name_last", "name_first", "name_middle",
  "phone", "address", "address_line_1", "address_line_2"
  )

# demographic variables
var_demographic <- c("age_disp", "gender", "date_of_birth", "race_ethnicity", "city", "state", "country")

# medical variables
var_medical <- c("height", "weight", "medical_category_historical", "past_medical_history", "medications", "allergies")

# emergency medical service variables
var_ems <- c("dispatch_date", "dispatch_day", "dispatch_mos", "dispatch_year", "service_name", "unit", "type_of_service_ihscene", "dispatch_priority", "response_mode", "transport_mode", "scene_lat", "scene_long", "reason_for_encounter", "medical_category", "chief_complaint", "drug_and_alcohol_indicators", "outcome_historical_all", "hpi", "scene_description", "base_region", "base_type", "alcohol", "drugs", "outcome_treated", "first_on_scene", "patient_activity_1", "patient_activity_2", "patient_activity_3")

# combined vector of all variable names
var_all <- c(var_id, var_demographic, var_medical, var_ems)

# order variables
df_ordered <- df_unique %>% select(!!!var_all)
```

## View Clean Data

```{r}
df_grca <- df_ordered %>% glimpse
```


# Outcome: HRI Case Definition

Define Heat Related Illness (HRI) case definition according to Noe 2013. 

1 hour outdoor heat exposure, with one of the following:

* HRI assessment or diagnosis
* HRI self-report
* Signs or symptoms of HRI without other etiology (e.g., URI, MVA)

## Screening HPI for Cases

Query the HPI variable for keywords, not case sensitive:

* Hot
* Heat
* Exhaust
* Dehydrat
* Hyponatrem

```{r}
df_screen <- df_grca %>% mutate(
  case_screen = ifelse(
    medical_category_historical == "Heat/Cold Exposure (2828)" 
    | str_detect(pattern = "(?i)hot|(?i)heat|(?i)exhaust|(?i)dehydrat|(?i)Hyponatrem", hpi),
    yes = 1, no = 2
    ) %>% form_yn()
)
```

Note, if any of the conditional variables (medical_category_historical, hpi) are missing, then case_screen is missing.

## Confirming HPI is consistent with HRI case definition

Medical professionals reviewed each case in the screening group to confirm whether case meets diagnosis criteria of HRI.

```{r}
case_confirm <- df_screen %>% 
  filter(case_screen == "Yes") %>% 
  select(prid, dispatch_date, medical_category, starts_with("patient_activity"), hpi)
case_confirm %>% glimpse
```


# Data Dictionary

This will be the data dictionary.


# Save Analytic Dataset

```{r}
saveRDS(df_screen, paste0(dir_data, "\\df_screen.rds"))
```
